---
title: "HW 3 Kaggle Walmart Data exploration and setup"
author: "Robert Kramer"
date: "November 19, 2015"
output: html_document
---


### Getting the data and putting it together
I put the data together unziped it and put it ./data

```{r,echo=FALSE, message=FALSE}
##setup
#setwd('./all_Robert_asus/prog_asus/R/HW/HW3')
library(readr)
library(ggplot2)
library(reshape)
library(stargazer)
library(Quandl)
library(ggthemes)

```
```{r,message=FALSE}
store <- read_csv("./data/stores.csv")
train <- read_csv("./data/train.csv")
test <- read_csv("./data/test.csv")
train <- read_csv("./data/train.csv")
features <- read_csv("./data/features.csv")
sampleSubmission <- read_csv("./data/sampleSubmission.csv")
```
Merged the data

```{r,message=FALSE}
## merge the store info 
train <- merge(train,store, all.x = T)
test <- merge(test,store, all.x = T)

## merge the features
train <- merge(train,features, all.x = T)
test <- merge(test,features, all.x = T)
```
I plotted the data to get a quick feel
```{r,message=FALSE}
qplot(x = Date,y = Weekly_Sales, data = train)
qplot(x = Type,y = Weekly_Sales, data = train)
```
Sorted it
```{r,message=FALSE}
## sort it by Date then Store then Dept
vars <- c('Store','Dept','Date')
train1 <- sort_df(train, vars)
```
In order to get a better understanding I first looked at 1 store.  Then 1 store 1 day. I looked at store 1 with all departments summed together.  I compared store 1 to total sales.  From this I assumed the stores where qualitatively the same. But...
```{r,message=FALSE}
store1 <- train1[train1$Store == 1,]
qplot(x = Date, y = Weekly_Sales, data = store1, colour=Dept)
summary(store1)

store1day1 <- store1[store1$Date==store1[[1,2]], ]
store1day1 <- sort_df(store1day1, 'Weekly_Sales')
summary(store1day1)

#trying to get aggregate of all dept by store and date
##all stores
test1 <- aggregate(store1$Weekly_Sales,list(store1$Date),sum)
test2 <- aggregate(train$Weekly_Sales,list(train$Date),sum)

Store1toAll <- cbind(test1,test2[2])
colVars <- c('date','store1_sales','total_sales')

names(Store1toAll) <- colVars
Store1toAll$date <- as.Date(Store1toAll$date)
Store1toAll <- data.frame(Store1toAll)

Store1toAll_plot <- ggplot(Store1toAll, aes(date,Store1toAll$store1_sales)) + geom_line(aes(color = "Store 1"))+
  geom_line(data = Store1toAll, aes(x= date,y=Store1toAll$total_sales/45, color = 'Mean of Total Sales'))+
    theme_economist() + scale_color_economist() +
    theme(legend.title = element_blank()) +
    labs(y= "Weekly Sales",title = ("Weekly sales total vs store 1"))
Store1toAll_plot
```
Then I took a look at the stores by Type and Type C is qualitatively different
```{r,message=FALSE}
A_stores <- subset(train1, train1$Type == 'A')
A_store_numbers <- levels(factor(A_stores$Store))
B_stores <- subset(train1, train1$Type == 'B')
B_store_numbers <- levels(factor(B_stores$Store))
C_stores <- subset(train1, train1$Type == 'C')  
C_store_numbers <- levels(factor(C_stores$Store))

## get a sense of the different types by weekly sales 
## aggregate then divide by the number of stores for average
A_stores_agg <- data.frame(aggregate(A_stores$Weekly_Sales,list(A_stores$Date),sum))
B_stores_agg <- data.frame(aggregate(B_stores$Weekly_Sales,list(B_stores$Date),sum))
C_stores_agg <- data.frame(aggregate(C_stores$Weekly_Sales,list(C_stores$Date),sum))

storeType_agg <- cbind(A_stores_agg,B_stores_agg[,2],C_stores_agg[,2])
names(storeType_agg) <- c('date','typeA','typeB','typeC')

plot <- ggplot(storeType_agg, aes(date,storeType_agg$typeA/length(A_store_numbers))) + geom_line(aes(color = "Type A"))+
  geom_line(data = storeType_agg, aes(x= date,y=storeType_agg$typeB/length(B_store_numbers), color = 'Type B'))+
  geom_line(data = storeType_agg, aes(x= date,y=storeType_agg$typeC/length(C_store_numbers), color = 'Type C'))+
  theme_economist() + scale_color_economist() +
  theme(legend.title = element_blank()) +
  labs(y= "Weekly Sales",title = ("Weekly sales mean by type"))
plot
```
I looked a little closer to make sure.
```{r,message=FALSE}
qplot(x = storeType_agg$date,y=storeType_agg$typeC,geom='line')

## type Cs look really wierd
store42 <- subset(train1,train1$Store == 42)
store42_agg <- aggregate(store42$Weekly_Sales,list(store42$Date),sum)
plot(store42_agg)
```
From this initial analysis I decided to pick a couple stores, one of type A and one of type B to begin an analysis without using the entire dataset.  I don't yet understand the type C's.
```{r,message=FALSE}
smallSet <- subset(train1, train1$Store == '1' | train1$Store == '3')
smallSet1 <- lm(Weekly_Sales ~. ,data=smallSet)
summary(smallSet1)
plot(smallSet1)


```